 Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation

### 2021-06-03 analysis/paper/shceof.Rmd
Pequeños detalles

The cEOF methodology is applied to SON geopotential height zonal anomalies at 50 and 200 hPa, representing a stratsopheric and upper troposhperic level respectively.
Table \@ref(tab:corr-ceof-splitted) shows the coefficient of determination between time series of the amplitude of each complex EOF (cEOF) across levels.
There's a high degree of correlation between the respective cEOF1 and cEOF2 at each level.
The spatial patterns of the 50 hPa and 200 hPa cEOFs are also similar (not shown).

@@

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
   # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.1, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "lr", size = 0.1, color = "gray30") +
  geom_line() +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", color = "black") +
  scale_x_datetime(NULL, date_labels = "%Y") +
  scale_y_continuous("EOF") +

  facet_grid(cEOF ~ part, scales = "free_y") +
  theme(panel.spacing = grid::unit(1.5, "lines"))  +
  tag_facets("cr") +
  panel_background
```

@@

(ref:eof1-regr-gh-cap) Regression coefficients of the real and imaginary part of the first cEOF on SON geopotential height anomalies for the 1979 -- 2019 period. These coefficients come from multiple linear regression involving the real and imaginary part of both cEFO2.

>>>>

The cEOF methodology is applied to SON geopotential height zonal anomalies south of 20\degree S at 50 and 200 hPa, representing a stratsopheric and upper troposhperic level respectively.
Table \@ref(tab:corr-ceof-splitted) shows the coefficient of determination between time series of the amplitude of each complex EOF (cEOF) across levels.
There's a high degree of correlation between the respective cEOF1 and cEOF2 at each level.
The spatial patterns of the 50 hPa and 200 hPa cEOFs are also similar (not shown).

@@

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
   # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.1, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "lr", size = 0.1, color = "gray30") +
  geom_line(aes(color = part)) +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", aes(color = part)) +
  scale_x_datetime(NULL, date_labels = "%Y") +
  scale_y_continuous("EOF") +
  scale_color_brewer(NULL, palette = "Dark2") +
  facet_grid(cEOF ~ part, scales = "free_y") +
  theme(panel.spacing = grid::unit(1.5, "lines"))  +
  tag_facets("cr") +
  panel_background
```

@@

(ref:eof1-regr-gh-cap) Regression coefficients of the real and imaginary part of the first cEOF on SON geopotential height anomalies for the 1979 -- 2019 period. 
These coefficients come from multiple linear regression involving the real and imaginary part of both cEFOs.
